\textbf{How to measure maximally precise radial velocities from a set of spectra:\\ The time-independent spectrum case}

### 2021-06-11 paper/eprv.tex
remove wrongish par

Another argument is as follows:
As long as the derivative $\dd f(v)/\dd v$ is close to orthogonal
(given the $C^{-1}$ metric) to the template $f(v)$, the derivative of
the log likelihood given in equation~(\ref{eq:dldv}) is equal to the
derivative of the cross-correlation $f(v)\T\cdot C^{-1}\cdot y$.
Therefore, whatever optimizes the likelihood will (nearly) optimize
the cross-correlation.

In actually performing this cross-correlation optimization,
investigators make some non-trivial choices.
The first is whether or not to use the inverse variance tensor $C$.
Many codes do not make any use of the uncertainty estimates at all in
the calculation of the cross correlation.
Not using the uncertainties (treating $C$ as the identity) is
equivalent to assuming that the pixel uncertainties don't vary much
from pixel to pixel across the spectrum.
If the cross correlation is optimized with $C$ set to the identity,
the optimum of that doesn't strictly optimize the likelihood function;
the differences depdends on the variations in the uncertainties.
Another choice is how finely to sample in radial velocity $v$, or how
to perform the optimization.
If the $v$ spacing is coarse (but not too coarse!), quadratic
interpolation can be used to find the optimum at high precision.
The velocity spacing need only be smaller than something like half the
spectral velocity resolution ($c\,\Delta\lambda / \lambda$).
This problem is similar to stellar centroiding methods in use in
imaging surveys (\CITE VAKILI).
Another choice faced by an investigator is what to do at the ends or
edges of the spectra or spectral segments used for the
cross-correlation.
If the edges are treated naively, it is possible to introduce unintended
dependencies that invalidate the requirement that the model-times-model term
(the $f(v)\T\cdot C^{-1}\cdot f(v)$ term) not depend on \RV\ $v$.

@@

\paragraph{Cross correlation with a binary mask}
The maximum-likelihood and cross-correlation methods only give
bound-saturating \RV\ estimates when the spectral model $f(v)$
respresents something similar to the true expecation of the spectrum
at velocity $v$.
And yet there is a paradox, because many world-class \RV\ pipelines
(like for instance \\CITE) perform cross-correlation with a
\emph{binary mask}.
This is not a good representation of the spectral expectation! So why
do these methods deliver high quality \RV\ estimates?
The answer is that the codes that cross-correlate with a binary mask
subsequently fit that cross-correlation with a Gaussian function.
It turns out that the operation ``cross-correlate the data with a
binary mask, followed by fitting a Gaussian to the cross-correlation
function'' is mathematically equivalent to ``smooth the binary mask
with a Gaussian, and then cross-correlate with the data''.
That is, when a pipeline cross-correlates with a binary mask and then
fits that cross-correlation function with a Gaussian, it is equivalent
to smoothing the binary mask with a best-fitting Gaussian function to
make a smooth, realistic $f(v)$ and then cross-correlating with that
realistic $f(v)$.
The fitting gives the flexibility to match the spectral resolution of
the template to the spectral resolution of the data, in the
approximation that the resolution is close to constant over the
spectral range in velocity units.
So in all cases of bound-saturating \RV\ measurements, the
\RV\ estimation is effectively a good approximation to
maximum-likelihood with a good spectral template.

>>>>

In actually performing this cross-correlation optimization,
investigators make some non-trivial choices.
The first is whether or not to use the inverse variance tensor $C^{-1}$.
Many codes do not make any use of uncertainty estimates at all in
the calculation of the cross correlation.
Not using the uncertainties (treating $C$ as the identity) is
equivalent to assuming that the pixel uncertainties don't vary much
from pixel to pixel across the spectrum.
If the cross correlation is optimized with $C$ set to the identity,
the optimum of that doesn't strictly optimize the likelihood function;
the differences depdends on the variations in the uncertainties.
Another choice is how finely to sample in radial velocity $v$, or how
to perform the optimization.
If the $v$ spacing is coarse (but not too coarse!), quadratic
interpolation can be used to find the optimum at high precision.
The velocity spacing need only be smaller than something like half the
spectral velocity resolution ($c\,\Delta\lambda / \lambda$).
This problem is similar to stellar centroiding methods in use in
imaging surveys (\CITE VAKILI).
Another choice faced by an investigator is what to do at the ends or
edges of the spectra or spectral segments used for the
cross-correlation.
If the edges are treated naively, it is possible to introduce unintended
dependencies that invalidate the requirement that the model-times-model term
(the $f(v)\T\cdot C^{-1}\cdot f(v)$ term) not depend on \RV\ $v$.

@@

\paragraph{Cross correlation with a binary mask}
The maximum-likelihood and cross-correlation methods only give
bound-saturating \RV\ estimates when the spectral model $f(v)$
respresents something similar to the true expecation of the spectrum
at velocity $v$.
And yet there is a paradox, because many world-class \RV\ pipelines
(like for instance \CITE) perform cross-correlation with a
\emph{binary mask}.
This is not a good representation of the spectral expectation! So why
do these methods deliver high quality \RV\ estimates?
The answer is that the codes that cross-correlate with a binary mask
subsequently fit that cross-correlation with a Gaussian function.
It turns out that the operation ``cross-correlate the data with a
binary mask, followed by fitting a Gaussian to the cross-correlation
function'' is mathematically equivalent to ``smooth the binary mask
with a Gaussian, and then cross-correlate with the data''.
That is, when a pipeline cross-correlates with a binary mask and then
fits that cross-correlation function with a Gaussian, it is equivalent
to smoothing the binary mask with a best-fitting Gaussian function to
make a smooth, realistic $f(v)$ and then cross-correlating with that
realistic $f(v)$.
The fitting gives the flexibility to match the spectral resolution of
the template to the spectral resolution of the data, in the
approximation that the resolution is close to constant over the
spectral range in velocity units.
So in all cases of bound-saturating \RV\ measurements, the
\RV\ estimation is effectively a good approximation to
maximum-likelihood with a good spectral template.
